---
title: "Knockoff_method_experiment1"
author: "Mark Milner"
date: "2025-01-19"
output: html_document
---

This experiment considers the number of samples (rows). A larger sample would suggest more info about the population and more precise estimates of model parameters.

```{r}
library(knockoff)
```
```{r}
library(doParallel)             # Load the package
numCores = detectCores()
print(numCores)
registerDoParallel(cores=numCores)
```

```{r}
fdp = function(selected, beta) {
  sum(beta[selected] == 0) / max(1, length(selected))
}
```

```{r}
power = function(selected, beta) {
  # True Positives: selected variables with non-zero coefficients
  TP = sum(beta[selected] != 0)
  
  # Total Relevant Variables
  T = sum(beta != 0)
  
  # Compute Power
  TP / T
}
```

## Data Setup 

```{r}
#set.seed(21)
```

```{r}
generate_synthetic_data <- function(n=1000, p=1000, relevant_count=100, amplitude=4.5, rho=0.25) {
  mu = rep(0,p)
  Sigma = toeplitz(rho^(0:(p-1)))
  X = matrix(rnorm(n*p),n) %*% chol(Sigma)
  nonzero = sample(p, relevant_count)
  beta = amplitude * (1:p %in% nonzero) / sqrt(n)
  y =  X %*% beta + rnorm(n)
  list(X = X, y = y, beta=beta)
}

```

# Implimentation

To implement this, we'll just iterate through a list of values. 

```{r}
n_s <- c(1000,2000,3000,4000,5000,6000,7000,8000,9000,10000)
fdp_s = c()
power_s = c()
for (n in n_s) {
  data <- generate_synthetic_data(n=n)
  X_n <- data$X
  y_n <- data$y
  beta_n <- data$beta
  result = knockoff.filter(X_n, y_n)
  current_fdp = fdp(result$selected,beta_n)
  current_power = power(result$selected,beta_n)
  fdp_s = c(fdp_s,current_fdp)
  power_s = c(power_s,current_power)
}
```








```{r}
fdp_s
```

```{r}
power_s
```

# Results visualisation 

## Simple FDR and power plots

We can start by simply plotting our n's on the x-axis and our performance metric on the y-axis 

```{r}
plot(n_s, fdp_s, type = "o", col = "red", pch = 16, lty = 1,
     xlab = "n", ylab = "FDP",
     main = "FDP vs number of rows (n)")

plot(n_s, power_s, type = "o", col = "blue", pch = 16, lty = 1,
     xlab = "n", ylab = "Power",
     main = "Power vs number of rows (n)")

```





